Higher-Order unification for free!

### 2024-04-10 paper/main.tex
rework 6 title

\noindent
We claim our compiler validates the following property:

@@

Since compilation moves problematic terms out of the sigh of \Ue{},
that procedure can only perform a partial occur check. For example the
unification problem $X \Ue f~Y$ cannot generate a cyclic substitution alone,
but should be disallowed if a $\linkStore$ contains a link like
$\linketaM{}{Y}{\lambda z.X_z}$: We don't know yet if $Y$ will feature
a lambda in head position, but we surely know it contains $X$, hence
$f~Y$ and that fails the occur check.
The procedure \elpiIn{occur-check-links} is in charge of ensuring that
each link does not represent a (suspended) unification problem doomed
to fail because of occur check. This check is needed in order to
guarantee \cref{prop:fidelity} (simulation fidelity).

@@

\begin{lemma}
  Let $\lambda x.t$ the \rhs of a \linketa, then $\wellb{t}$.
  \label{lemma:unif-eta-aux}
\end{lemma}

@@

\begin{lemma}[Fidelity with \linketa]
  The introduction 
  of \linketa guarantees \fullref{prop:fidelity} 
  \label{lemma:fidelity-maybeeta}
\end{lemma}

@@

\section{Enforcing inv.: \NoCaseChange{\nameref{inv:uvaarity}}}
\label{sec:invariant1}

We report here the problem given in \cref{sec:basic-comp-limitations} where
$X$ is used with two different arities and the output of the compilation does
not respect \cref{inv:map-bijective}. Moreover, merging the two mappings\noteD{TODO: da cambiare 488 in tex}
for $s$ would break \fullref{inv:uvaarity}.
In this section we explain how to replace the duplicate mapping with
\linketa in order to restore the invariants.
% make test ONLY=7001 TEX=tex
\printAlll
  {{{\lambda x.\lambda y.(X\appsep y\appsep x),\lambda x.\lambda y.x},
    {\lambda x.(f\appsep (X\appsep x)\appsep x),Y}}}
  {{{A,\lambda x.\lambda y.x},
    {D,F}}}
  {{{X,E,1},
    {Y,F,0},
    {X,C,2}}}
  {{{\eta,,D,\lambda x.(f\appsep E_{x}\appsep x)},
    {\eta,,A,\lambda x.B_{x}},
    {\eta,x,B_{x},\lambda y.C_{y x}}}}

>>>>

\noindent
We write $\wellb(t)$ as a short for $\wellb(\{t\})$.
We claim our compiler validates the following property:

@@

Since compilation moves problematic terms out of the sigh of \Ue{},
that procedure can only perform a partial occur check. For example the
unification problem $X \Ue f~Y$ cannot generate a cyclic substitution alone,
but should be disallowed if a $\linkStore$ contains a link like
$\linketaM{}{Y}{\lambda z.X_z}$: We don't know yet if $Y$ will feature
a lambda in head position, but we surely know it contains $X$, hence
$f~Y$ and that fails the occur check.
The procedure \elpiIn{occur-check-links} is in charge of ensuring that
each link does not represent a (suspended) unification problem doomed
to fail because of occur check. This check is needed in order to
guarantee \fullref{prop:fidelity}.

@@

\begin{lemma}
  Let $\lambda x.t$ the \rhs of a \linketa, then $\wellb(t)$.
  \label{lemma:unif-eta-aux}
\end{lemma}

@@

\begin{theorem}[Fidelity in \maybeeta]
  The introduction 
  of \linketa guarantees \fullref{prop:fidelity} 
  \label{lemma:fidelity-maybeeta}
\end{theorem}

@@

%\section{Enforcing inv.: \NoCaseChange{\nameref{inv:uvaarity}}}
\section{Making \mapStore a bijection}
\label{sec:invariant1}
\todo{l'esempio è sparito}
We report here the problem given in \cref{sec:basic-comp-limitations} where
$X$ is used with two different arities and the output of the compilation does
not respect \cref{inv:map-bijective}. Moreover, merging the two mappings
for $s$ would break \fullref{inv:uvaarity}.
In this section we explain how to replace the duplicate mapping with
\linketa in order to restore the invariants.
% make test ONLY=7001 TEX=tex
\printAlll
  {{{\lambda x.\lambda y.(X\appsep y\appsep x),\lambda x.\lambda y.x},
    {\lambda x.(f\appsep (X\appsep x)\appsep x),Y}}}
  {{{A,\lambda x.\lambda y.x},
    {D,F}}}
  {{{X,E,1},
    {Y,F,0},
    {X,C,2}}}
  {{{\eta,,D,\lambda x.(f\appsep E_{x}\appsep x)},
    {\eta,,A,\lambda x.B_{x}},
    {\eta,x,B_{x},\lambda y.C_{y x}}}}
